---
title: "Testing and Developing diplomix"
output: html_notebook
---


## Simulation Function

We need to simulate some data to see how well we can recover this stuff.

```{r}
library(tidyverse)
library(tetradiplomix)
```

We pass it a vector of ancestral allele freqs and an F value and then we use the
Nicholson model to simulate new values.
```{r}
#' Simulate tetraploid data that is a mixture of two diverging loci
#' 
#' @param a vector an ancestral frequencies/weights of alleles
#' @param F drift parameter
#' @param N number of individuals
simdiplomix <- function(a, F, N) {
  
  # simulate the zeds for each indvidual randomly
  zm <- zed_matrix()
  zim <- zm[sample(1:nrow(zm), size = N, replace = TRUE), ]
  
  # simulate the allele freqs
  an <- a / sum(a)
  
  a0 <- rgamma(n = length(a), shape = an * (1 - F) / F, scale = 1)
  a0 <- a0 / sum(a0)
  
  a1 <- rgamma(n = length(a), shape = an * (1 - F) / F, scale = 1)
  a1 <- a1 / sum(a1)
  
  # sample them into the individuals
  G <- matrix(0, nrow = N, ncol = 4)
  
  n0 <- sum(zim == 0)
  n1 <- sum(zim == 1)
  
  G[zim == 0] <- sample(x = 0:(length(a) - 1), size = n0, replace = TRUE, prob = a0)
  G[zim == 1] <- sample(x = 0:(length(a) - 1), size = n1, replace = TRUE, prob = a1)
  
  list(G = G, p = rbind(a0, a1), zim = zim)
}
```

Now that we have that, we can easily simulate 96 individuals at a tetraploid locus
that behaves like two diploid loci with Fst of about 2 * 0.03 and 5 alleles
```{r}
Gsim <- simdiplomix(a = 5:1, F = 0.03, N = 96)
```



## Testing update_zim

I want to simulate some things that are very diverged and check to make sure
that we have things mostly right...(hmmm...)

```{r}
test <- simdiplomix(a = 5:1, F = 0.4, N = 96)
p <- rbind(test$a0, test$a1)
zm <- zed_matrix()
zim <- matrix(0, nrow = nrow(test$zim), ncol = ncol(test$zim))


boing <- update_zim(test$G, zim, zm, p)

# note that most of these are TRUE, so I think we have it right.
test$zim == boing
```


## Now, run diplomix through its paces

I have completed the coding for the function `diplomix()` and it seems to be 
working well.  Now I am going to do a little experiment: I will simulate
data at different Fst values and then run it through `diplomix()` and assess
the trace of the Fst value.
```{r}
Ftrue <- c(0.01, 0.02, 0.03, 0.05, 0.1, 0.2) 
names(Ftrue) <- Ftrue

simout <- lapply(Ftrue, function(x) {
  lapply(1:10, function(y) {
    inp <- simdiplomix(a = 5:1, F = x, N = 96)
    out <- diplomix(inp$G, 500)
    tibble(rep = y, iter = 1:500, fst_trace = fst_from_ptrace(out$ptrace))
  }) %>%
    bind_rows()
}) %>% 
  bind_rows(.id = "sim_F")
```

Now we can make a massive fact plot of all that:
```{r}
ggplot(simout, aes(x = iter, y = fst_trace, group = rep)) +
  geom_line() +
  facet_wrap(~ sim_F, ncol = 3)
```
